---
title: "Crime in Minneapolis"
author: "ETC5513 Group 6"
date: "30/05/2020"
output: html_document
---

Minneapolis has the highest violent crime rate in the US. 

Research Questions (If anyone thinks of more, please add!!) 
- Where is crime most common in Minneapolis?
Heat map
- What is the ethnicity of those who are commmitting the crimes? 
Major crimes/minor crimes
- Where is force used?
Heat Map 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(rgdal)
```

```{r}
crime_incidents <- read_csv("data/Police_Incidents_Last_2Years.csv") 
```

```{r}
shp <- rgdal::readOGR(dsn = file.path(paste(getwd(),"/data/Police_Incidents_Last_2Years-shp", sep=""), layer = "Police_Incidents_Last_2Years.shp") , stringsAsFactors = FALSE)

summary(shp@data)
```

```{r}
police_use_of_force <- read_csv("data/Police_Use_of_Force.csv")
```

```{r}
police_stop_data <- read_csv("data/Police_Stop_Data.csv")
```

Minneapolis has the highest violent crime rate in the US. 
Research Questions (If anyone thinks of more, please add!!) 
- Where is crime most common in Minneapolis?
Heat map
- What is the ethnicity of those who are commmitting the crimes? 
Major crimes/minor crimes
- Where is force used?
Heat Map 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


What; no. different crime offense?
```{r}
crime_incidents %>%
  group_by(offense) %>%
  summarise(unique(offense))
```

Grouping crime offense 
```{r clean-crime}
clean_crime <- crime_incidents %>%
  mutate(offense = case_when(
    offense %in% c("ARSON") ~ "Arson",
    offense %in% c("ASLT1",
                   "ASLT2",
                   "ASLT3",
                   "ASLT4") ~ "Assault",
    # Combining automobile theft, bike theft, coin-operated device theft, gas-station drive off, online theft, petty theft, pocket picking, scrapping-recycling theft, theft from a building, theft from a motor vehicle, theft from a person, other theft, theft by swindle, theft of motor vehicle parts
    offense %in% c("AUTOTH",
                   "MVTHFT",
                   "BIKETF",
                   "COINOP",
                   "NOPAY",
                   "ONLTHT",
                   "PETIT",
                   "POCKET",
                   "SCRAP",
                   "TBLDG",
                   "TFMV",
                   "TFPER",
                   "THEFT",
                   "THFTSW",
                   "TMVP") ~ "Theft",
    # Combining burglary of business and burglary of dwelling
    offense %in% c("BURGB",
                   "BURGD") ~ "Burglary",
    offense %in% c("ROBBIZ",
                   "ROBPAG",
                   "ROBPER") ~ "Robbery",
    offense %in% c("SHOPLF") ~ "Shoplifting",
    offense %in% c("COMPUT") ~ "Computer Hacking",
    offense %in% c("CSCR") ~ "Rape",
    # Combing first, second and third degree domestic assult with domestic assault strangulation
    offense %in% c("DASLT1",
                   "DASLT2",
                   "DASLT3",
                   "DASTR") ~ "Domestic Assault",
    offense %in% c("DISARM") ~ "Disarm a police officer",
    offense %in% c("LOOT") ~ "Looting",
    offense %in% c("MURDR") ~ "Murder"))
clean_crime
```

no. of crimes in each group 
```{r number-crimes}
number_crimes <- clean_crime %>%
  group_by(offense) %>%
  summarise(Incidence = n())

number_crimes
```

no. of crimes; past for last 2 years
```{r plot-number-crimes}
ggplot(number_crimes,
       aes(x = offense,
           y = Incidence)) +
  geom_col() +
  xlab("Type of Offense") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3))
```




# Police use of force data

What; type of police forcetype?
```{r police-use-of-force-clean}
police_use_of_force %>%
  group_by(ForceType) %>%
  summarise(unique(ForceType))
```

number of times; force types used by police
```{r}
police_use_of_force %>% 
  group_by(ForceType) %>% 
  count()
```

what; no. of times force types used against each race
```{r}
police_use_of_force %>% 
  group_by(ForceType, Race) %>% 
  count()
```

No. of incidence; force used; per race
```{r}
police_use_of_force %>% 
  group_by(Race) %>% 
  count()
```



Use of force per race faceted by Force Type
```{r}
use_of_force_bar <- police_use_of_force %>% 
  filter(!is.na(Race),
         ForceType != "Bodily Force") %>% # see other types; many Bodily Force incidence; extend the ylimit; so remove
  ggplot() +
  geom_bar(aes(x = Race)) +
  facet_wrap(~ `ForceType`) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3)) +
  coord_flip()

use_of_force_bar
```

```{r bodily-force-pie-chart}
use_of_force_bar +
  coord_polar()
```


# attempting to do maps 

```{r}
library(ggmap)
library(leaflet)

make_bbox <- ggmap::make_bbox(lon = police_use_of_force$X,
                     lat = police_use_of_force$y, f = 0.9)  

minneapolis_map <- leaflet() %>%
  # Base group
  leaflet::addProviderTiles(provider = providers$Esri.WorldImagery, group="Satellite") %>%
  # Add reciever location data
  addCircles(
    lng = police_use_of_force$X, 
    lat = police_use_of_force$Y, 
    fill=TRUE, color="yellow",
    weight= 2,
    radius = 2,
    stroke = FALSE, fillOpacity = 0.8)

minneapolis_map

```














